---
title: "AminoAcid_LTRTracker"
author: "NathanGLawler"
date: "`r Sys.Date()`"
output: html_document
---

#Library Installation
```{r}
#Install non-standard required packages
# devtools::install_github("jwist/fusion")
# install.packages("tidyverse")
# install.packages("gridExtra")
# install.packages("cowplot")
# devtools::install_github("tkimhofer/metabom8")
```

#Load Libraries
```{r, message=FALSE}
#Load required libraries

#### load packages
package_list <- c('tidyverse', 'janitor', 'qualV', 'kableExtra', 'fusion',
                  'readxl', 'ggforce','broom')
for(idx_package in package_list){
  if(length(which(row.names(installed.packages()) == idx_package)) > 0){
    suppressMessages(require(package = idx_package,
                             character.only = TRUE))
  } else {
    paste0(idx_package, " is not installed. Please install ", idx_package, " before continuing.") %>% print()
  }
}

```

##Read in file configuration for Amino Acids from a centrally located google sheet. The sheet can be altered by a user to include or not include metabolites to review and past AA cohorts.

```{r}

pathToFig <- file.path(Sys.getenv()['PROJECTS'], "LTRCheckR", "AA")

metConfig <- read.csv2("https://docs.google.com/spreadsheets/d/e/2PACX-1vRPDk-U3YlaSvCct-lYcBmB5qQekXdY9q71726yxzkbrX-vFZLQ6V29y_aLroadUTmf4EPr8gz6OzSM/pub?gid=190423913&single=true&output=csv", header = TRUE, sep = ",")

fileConfig <- read.csv2("https://docs.google.com/spreadsheets/d/e/2PACX-1vRPDk-U3YlaSvCct-lYcBmB5qQekXdY9q71726yxzkbrX-vFZLQ6V29y_aLroadUTmf4EPr8gz6OzSM/pub?gid=631312894&single=true&output=csv", header = TRUE, sep = ",")



```


#Load Config Files
```{r}
#Read config file
# metConfig <- readxl::read_excel("./config/configFile.xlsx", sheet="metabolites")
# fileConfig <- readxl::read_excel("./config/configFile.xlsx", sheet="dataFiles")

#Extract list of targeted metabolites to be included (incl. Internal Standards)
metISTDList <- metConfig %>% dplyr::filter(toInclude==TRUE)
metISTDList <- dplyr::select(metISTDList, 1)

#Extract list of targeted metabolites to be reviewed (excl. Internal Standards)
metList <- metConfig %>% dplyr::filter(toReview==TRUE)
metList <- dplyr::select(metList, 1)

#Extract list of data files to review
fileList <- fileConfig %>% dplyr::filter(toReview==TRUE)
fileList <- dplyr::select(fileList, 1)
```



#Load All Data Elements
```{r}
ltrQuant <- NULL
# ltrArea <- NULL
sampleCohort <- NULL

# dataDir <- "./data/"
# dataFiles <- list.files(path=dataDir, full.names=FALSE)

for(i in 1:nrow(fileList)){
# for(fileName in dataFiles){
  #Load data element file
  load(fileList$fileName[i])
  # load(paste(dataDir, fileName, sep=""))

  #Extract all LTR data from project daE file
  sampleType <- fusion::getData(daAA, type = c("ltr"), replicates = "all")

  #LTR data file name & cohort
  sampleID <- sampleType@obsDescr[[1]]$AnalysisName
 
  # cohort <- sapply(strsplit(fileName, split="_"), getElement, 2)
  cohort <- sapply(strsplit(fileList$fileName[i], split="_"), getElement, 2)
  sc <- data.frame(sampleID, cohort)

  #LTR Quantity
  quant <- dplyr::as_tibble(sampleType@.Data) %>% as.data.frame()
  colnames(quant) <- unname(sapply(colnames(quant), function(x) gsub(" [IS]", "", x, 
                                                                     fixed = TRUE)))
  quant[quant == "0"] <- NA
  quant <- quant %>% dplyr::select(any_of(metList$metaboliteName))

  #Area of LTR
  # area <- do.call("cbind", (lapply(sampleType@obsDescr, function(x) x$Area))) %>% as.data.frame()
  # colnames(area) <- sampleType@varName
  # colnames(area) <- unname(sapply(colnames(area), function(x) gsub(" [IS]", "", x, fixed = TRUE)))
  # area <- area %>% dplyr::select(metList$metaboliteName)
  
  #Check for same colnames before rbind
  # cols<- intersect(colnames(ltrQuant), colnames(quant))
  
  #Join all LTR data frames
  ltrQuant <- bind_rows(ltrQuant, quant)
  # ltrArea <- rbind(ltrArea, area)
  sampleCohort <- rbind(sampleCohort, sc)
  
  #Remove unused variables
  # rm(daAA, sampleType, sampleID, cohort, sc, quant, area)
  rm(daAA, sampleType, sampleID, cohort, sc, quant)
}

#Provide unique number for each sample ID
sampleCohort$idx <- 1:nrow(sampleCohort)

#Remove unused variables
# rm(i)
rm(dataFiles, fileName)
```

#NA% Plot Generation (LTR Quant)
```{r}
#Identify the number of NA values present for each metabolite
plotData_NA <- apply(ltrQuant, 2, function(x){
  c("TRUE"=length(which(is.na(x))), "FALSE"=length(which(!is.na(x))))
})
plotData_NA <- as.data.frame(plotData_NA, stringsAsFactors=FALSE)
plotData_NA$missingValues = rownames(plotData_NA)

#Convert the data matrix to long format
plotData_NA <- reshape2::melt(plotData_NA, id.vars = "missingValues")

#Plot graph
ggplot(plotData_NA, aes(variable, value, fill=missingValues))+
  geom_bar(stat="identity")+
  coord_flip()+theme_bw()+
  scale_fill_manual(values=c("light green", "red"))+
  labs(y=NULL, x=NULL)+
  scale_y_continuous(sec.axis = sec_axis(~./nrow(ltrQuant)*100, name='%'), name='n')
```

#Data Cleaning (LTR Quant)
```{r}
# Remove metabolites with NA values in its rows
# ltrQuant <- ltrQuant[ , colSums(is.na(ltrQuant))  > 20]
ltrQuant <- ltrQuant[ , (colMeans(is.na(ltrQuant)) * 100) < 50]

ltrQuant <- apply(ltrQuant, 2, function(x){
  as.numeric(x)
  }) %>% as.data.frame()

#Merge two data frames together
bootLTRQ <- cbind(ltrQuant, sampleCohort)

# #Set data in long format
bootLTRQ <- melt(bootLTRQ, id.vars = colnames(sampleCohort))

```

#Bootstrap Confidence Interval (CI)
```{r}

# bootSample <- list()
# 
# for(i in 1:1000){
#   fi <- sample(nrow(ltrQuant), (0.80*nrow(ltrQuant)))
#   metStatsVar <- bootLTRQ %>% dplyr::group_by(variable) %>% 
#     dplyr::summarise(median = median(value[fi], na.rm = TRUE), sd = sd(value, na.rm = TRUE), 
#                      sd2 = sd(2*value, na.rm = TRUE), minv = min(value, na.rm = TRUE), maxv = max(value, na.rm = TRUE), se = sd/sqrt(nrow(ltrQuant)))
#   bootSample[[i]] <- metStatsVar$median
# }
# bootSample <- do.call("cbind", bootSample) %>% as.data.frame()
# lowerBound <- apply(bootSample, 1, function(x) quantile(x, 0.025))
# upperBound <- apply(bootSample, 1, function(x) quantile(x, 0.975))
# midPoint <- rowMeans(bootSample)
# 
# 
# metStatsVar <- cbind(metStatsVar, lowerBound, midPoint, upperBound) %>% as.data.frame()
# 
# metStatsVar <- metStatsVar %>% mutate_if(is.numeric, round, digits = 3)
# htmlTable::htmlTable(metStatsVar)
# 
# #Remove unused variables
# rm(i, fi, lowerBound, upperBound, midPoint)
```

#LTR Plot Generation
```{r}


allLTR <- cbind(sampleCohort, ltrQuant) %>% select(-idx)

##Flip data to long format
LTRLong <- allLTR %>% 
  pivot_longer(cols = "5-hydroxyindole acetic acid":"SIL Tryptophan-d5_neg",
                 names_to = "key", 
                 values_to = "value") %>% 
  rowid_to_column(var = "idx")
 
LTRLong$key <- as.factor(LTRLong$key)

ltrSER <- LTRLong %>% 
  ggplot(aes(x = idx, y = value, colour = cohort)) +
  geom_point() +
  facet_wrap_paginate(~ key, ncol = 4, nrow = 5, page = 1, scales = "free_y") 
  n <- n_pages(ltrSER)

pdf(file = file.path(pathToFig,'Serum.pdf'), paper= 'A4', w= 210/25.4, 297/25.4)
for(i in 1:n){
    print(ltrSER + facet_wrap_paginate(~ key, ncol = 3, nrow = 5, page = i, scales = "free_y"))
}
dev.off()




# plotData_LTRQ <- list()
# 
# for(i in 1:ncol(ltrQuant)){
#   plot_data <- dplyr::as_tibble(cbind(ltrQuant[, i], sampleCohort$sampleID, sampleCohort$cohort, sampleCohort$idx)) %>%
#     mutate_at(c("V1", "V4"), as.numeric)
#     colnames(plot_data) <- c("value", "sampleID", "cohort", "idx")
#     plot_data$metabolite <- colnames(ltrQuant[i])
#     
#   # plot_data <- plot_data %>%
#   #   summarise(metabolite = metabolite, sampleID = sampleID, idx = idx, cohort = cohort, value = value, 
#   #             lowerCI = metStatsVar$lowerBound[i], mean = metStatsVar$midPoint[i], 
#   #             upperCI = metStatsVar$upperBound[i], median = metStatsVar$median[i], 
#   #             sd = metStatsVar$sd[i], sd2 = metStatsVar$sd2[i], se = metStatsVar$se[i]) %>%
#   #   as.data.frame()
#   
#   pdLTRQ <- ggplot(plot_data, aes(idx, value, color=factor(cohort)))+
#         # geom_hline(aes(yintercept = lowerCI), linetype="dotted", colour="BLACK")+
#         # geom_hline(aes(yintercept = upperCI), linetype="dotted", colour="BLACK")+
#         # geom_hline(aes(yintercept = mean - sd), linetype="dotted", colour="RED")+
#         # geom_hline(aes(yintercept = mean + sd), linetype="dotted", colour="RED")+
#         # geom_hline(aes(yintercept = mean - sd2), linetype="dotted", colour="GREEN")+
#         # geom_hline(aes(yintercept = mean + sd2), linetype="dotted", colour="GREEN")+
#         geom_point(shape=1)+
#         theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
#         panel.background = element_blank(), axis.line = element_line(colour="BLACK"))+
#         facet_wrap(.~metabolite, nrow = 10, ncol = 6, scales = "free_y")+
#         scale_y_continuous(trans = "log10", name = "Concentration uM")+
#         # ylim(c(min(plot_data$value - 0.2), max(plot_data$value + 0.2))) +
#         labs(x = "Run Order")
#   
#   plotData_LTRQ[[i]] <- pdLTRQ
# }
# 
# #Remove unused variables
# rm(i)

```

#Calculate RSD between cohorts on LTR values for each metabolite
```{r}
df <- ltrQuant

gather(df, factor_key=TRUE) %>% 
  group_by(key) %>% 
  summarise(mean=mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), rsd = (sd/mean)*100, min=min(value), max=max(value)) -> statsLTR

statsLTR$colour <- ifelse(statsLTR$rsd >25, "firebrick", "dodgerblue")

# test <- sapply(df, function(x) c( "Stand dev" = sd(x, na.rm = TRUE), 
#                          "Mean"= mean(x,na.rm=TRUE),
#                          "n" = length(x),
#                          "Median" = median(x, na.rm = TRUE),
#                          "CoeffofVariation" = sd(x)/mean(x,na.rm=TRUE)*100,
#                          "Minimum" = min(x),
#                          "Maximun" = max(x)))

rsdThreshold <- 25 ##User input required here

plotRSD <- ggplot(statsLTR, aes(x = key, y = rsd, color = as.factor(colour))) +
  geom_point() +
  geom_hline(yintercept = 25, linetype = "dashed")+
  geom_label_repel(data = statsLTR %>% filter(rsd>rsdThreshold),
             aes(label=key)) +
  scale_color_manual(values=c("dodgerblue", "firebrick")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position = "none") +
  labs(x = "Metabolite", y = "rsd%") +
  lims(y=c(0,max(statsLTR$rsd+10)))
  
  





```

